---
title: "Mini Project 3"
author: "Chelsey Chen, Yuri Furukawa"
date: "April 15, 2018"
output:
  html_document:
    code_folding: show
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
#devtools::install_github("beanumber/macleish")
library(macleish)
packageVersion("tidyverse")
names(macleish_layers)
library(RColorBrewer)
library(leaflet)
library(sf)
library(units)

```


##Data Wrangling 
```{r, message = FALSE, warning = FALSE}
# Calculate total length for each trail in Macleish Field Station
trails <- macleish_layers[["trails"]]
trails <- trails %>%
  mutate(LENGTH = st_length(geometry)) %>%
  group_by(name) %>%
  summarize(num_segments = n(), 
            total_length = sum(LENGTH))


# Join the 3-meter Contours layer to macleish trails layers and make that as a new dataframe
my_data <- trails %>%
  st_join(macleish_layers[["contours_3m"]], join = st_intersects) %>%
  select(-num_segments,-OBJECTID, -Shape_Leng, -Shape_Area, -ELEV_FT, -INDEX_ARC, -SHAPE_LEN) %>%
  group_by(name,total_length) %>%
  summarize(change_elevation = max(ELEV_M)-min(ELEV_M)) %>%
  mutate(score_elevation = ifelse(change_elevation %in% 0:30, 0,
                           ifelse(change_elevation %in% 31:60, 1,2)),
         score_length=ifelse(total_length < 750 * with(ud_units,m), 0,
                      ifelse(total_length >= 750* with(ud_units,m) && 
                             total_length < 1500* with(ud_units,m), 1, 2)),
         total_score = score_elevation + score_length,
         difficulty = ifelse(total_score %in% 0:1, "Easy",
                      ifelse(total_score %in% 2:3, "Moderate", "Difficult")))
```
```{r}
# Reorder the trail's difficulty factors
my_data$difficulty <- factor(my_data$difficulty, levels = c("Easy","Moderate","Difficult"))
```

##Data Visualizations
```{r, message = FALSE, warning = FALSE}
# Create a custom color palette for the trail's difficulty
trail_pal <- colorFactor("viridis", 
                  my_data$difficulty)

# Create a leaflet interacting map
leaflet() %>%
  addTiles() %>%
  addPolylines(data = my_data,
               popup = ~name,
               group = ~difficulty,
               color = ~trail_pal(difficulty),
               fillOpacity = 0.9, weight = 3.5) %>%
  addCircles(data = st_cast(macleish_layers[["landmarks"]],"POINT"),
             popup = ~Label) %>%
  #addPolygons(data = macleish_layers[["streams"]]) %>%
  addLegend(position = "bottomright",
            pal = trail_pal,
            values = my_data$difficulty, 
            title = "Difficulty")
```

## Data manipulation
MacLeish Field Station has roughly four miles of built trail for use by field station visitors. In this mini project, we classified these trails as “Easy,” “Moderate,” or “Difficult” based on their length and change in elevation. This classification helps visitors to perceive which trail is relatively easier to walk and therefore makes a wiser decision on the visiting route.
First, we extracted the trail layer under the macleish_layers spatial data. Since many of the trails have multiple segments, we calculated the total length of each trail. We then joined this trail layer to the MacLeish 3-meter contours data ("contours_3m") by intersections. After joining these two spatial layers together and calling the returning dataframe as "my_data", we removed some of the variables that we do not care about for this project; in other words, we only kept the name of the trail ("name"), total length of the trail ("total_length"), elevation in meter ("ELEV_M"), and the geometry columns. We then added a column of the maximum change in elevation for each trail by calculating the difference between higheset and lowest elevation.
Second, we assigned a score ("score_length") based on the total length of each trail. If a trail's total length falls in the range of 0 and 750 meters, that trail would score 0; if its length falls in the range of 750 to 1500 meters, a trail would score 1; if its length is over 750 meters, it would score 2. We then assigned a score ("score_elevation") based on the maximum change in elevation for each trail. If a trail has a maximum change in elevation between 0 and 30 meters, the trail would score 0, and it would score 1 for 31-60 meters range, and 2 for >60 meters range. After calculating the scores based on length and change in elevation, we combined these two scores into an overall score for the difficulty for using that trail. By calculating this overall score of difficulty, we categorized all the MacLeish trails into three categories. If a trail scores 0 or 1, it is an "Easy" trail; if it scores 2 or 3, it is a "Moderate" trail; if it scores 4, it is a "Difficult" trail.

## Leaflet Map
After the data manipulation, we created an interacting leaflet map that shows every MacLeish trail colored by its difficulty. We first created a color palette using the difficulty category as the values. We then created a leaflet map and every trail is colored by its difficulty. Purple trails represents the easy ones, green represents the moderate ones, and the yellow represents the difficult ones.



## Not execute
```{r}
#-----------------------------
  proj4_aea <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

road_buffer <- macleish_layers[["trails"]] %>%
  st_transform(proj4_aea) %>%
  st_buffer(dist = 10) %>%
  st_transform(4326)

leaflet() %>%
  addTiles() %>%
  addPolylines(data = macleish_layers[["trails"]], 
               weight = 1, color = "black") %>%
  addPolygons(data = road_buffer) 
```

